---
title: "Course Project 1"
author: "Vamshidhar"
date: "October 26, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#My first markdown, Course project 1:  
####Loading required pacakges:
loading packages that are helpfull in our analysis.
```{r}
suppressMessages(library(data.table))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
```

## Loading and preprocessing the data:
Lets start our analysis by downlaoding, uploading and discovering the data.
```{r cacahe = TRUE}
link <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(link, destfile = "data.zip")
unzip("data.zip")
amd <- fread("activity.csv", header = TRUE ,na.strings = "NA")
amd <- tbl_df(amd)
summary(amd)
```

## What is mean total number of steps per day?
#### Caliculating total no of steps per each day:
```{r }
tot_stepsinday <- aggregate(steps ~ date, data = amd, sum, na.rm = TRUE)
```

####Ploting histogram: 

```{r }
ggplot(data = tot_stepsinday, aes(x = tot_stepsinday$steps)) +
        geom_histogram(aes(fill = ..count..), bins = 15) +
        scale_fill_gradient("Count", low = "yellow", high = "blue") +
        xlab( "Total no of steps per day") +
        ggtitle("Histogram of total number of steps taken per day")

```

####Caliculating Mean and Median:

```{r }
Mean <- round(mean(tot_stepsinday$steps))
Median <- median(tot_stepsinday$steps)
```
**Observation :** Mean and Median of the number of steps taken per day was caliculated in the above code, they are found to be Mean = `r sprintf("%.0f",Mean)` and Median = `r Median`

## What is the average daily activity pattern?

```{r }
avg_pattrn <- aggregate(steps ~ interval, data = amd, mean, na.rm = TRUE)
d <- which(avg_pattrn$steps == max(avg_pattrn$steps))
max <- avg_pattrn[d,1]
ggplot(avg_pattrn, aes(x = interval, y = steps,label = 
                               paste(avg_pattrn[d,1], "th interval",sep = ""))) +
        geom_line() +
        geom_text(data = avg_pattrn[d,],vjust=0, colour="red")
```

**Observation :**interval that contains the maximum number of steps is `r max`  

## Imputing missing values:
```{r }
miss <- sum(!complete.cases(amd))
```
Total number of missing values in the dataset are `r miss`

## Imputation of missing values:
####Filling missing values with its *mean*
* Strategy used to impute missing values here is simple by *mean*
```{r }
imputedamd <- amd
imputedamd$steps[imputedamd$steps %in% NA] <- mean(imputedamd$steps, na.rm = TRUE)
colSums(is.na(imputedamd))
```

```{r }
tot_stepsinday_imputed <- aggregate(steps ~ date, data = amd, sum, na.rm = TRUE)
ggplot(data = tot_stepsinday_imputed, aes(x = steps)) +
        geom_histogram(aes(fill = ..count..), bins = 15) +
        scale_fill_gradient("Count", low = "yellow", high = "blue")+
        xlab( "Total no of steps per day") +
        ggtitle("Histogram of total number of steps taken per day in imputed dataset")
```

**Observation :** There is no significant change in histograms before and after imputing missing valyes:

```{r }
imputed_Mean <- round(mean(tot_stepsinday_imputed$steps))
imputed_Median <- median(tot_stepsinday_imputed$steps)
```

Parameter      |        Mean                    | Median
-------------- | -----------------------------  | ------------------
actual data    |`r sprintf("%.0f",Mean)       ` | `r Median`
imputed data   |`r sprintf("%.0f",imputed_Mean)`|`r imputed_Median`

**Observation :** As observed above, Mean and Median values does not change even after imputation of missing values.

## Are there differences in activity patterns between weekdays and weekends?
#### using imputed dataset here:
```{r cache = TRUE}
imputedamd$date <- as.Date(strptime(imputedamd$date, format="%Y-%m-%d"))
imputedamd$day <- weekdays(imputedamd$date)
for (i in 1:nrow(imputedamd)) {
        if (imputedamd[i,]$day %in% c("Saturday","Sunday")) {
                imputedamd[i,]$day<-"weekend"
        }
        else{
                imputedamd[i,]$day<-"weekday"
        }
}
stepsperday <- aggregate(imputedamd$steps ~ imputedamd$interval + imputedamd$day, imputedamd, mean)
```

####Ploting weekend vs weekday to compare weekend and weekday pattern

```{r }
names(stepsperday ) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsperday , type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```

#*********End of the project************